---
title: "EDA: Integrated Broad/Sanger Institute CRISPR-Cas9 Data"
output: html_notebook
---

# Summary
This notebook explores the Integrated CRISPR-Cas9 dataset from Pacini, C., Dempster, J. M., Boyle, I., Gonçalves, E., Najgebauer, H., Karakoc, E., ... & Iorio, F. (2021). Integrated cross-study datasets of genetic dependencies in cancer. Nature communications, 12(1), 1-14.

The reason why we're interested in this dataset is because it has been corrected by copy number variation, has been batch corrected for each study and multiple variables, and has greater statistical power due to the combined dataset. 

```{r}
rm(list=ls())
```

# 1. Load libraries
```{r}
# General libraries
library(tidyr)
library(readr)
library(stringr)
library(reticulate)
library(tidyverse)
library(dplyr)
library(reshape2)
library(readxl)

# Visualization
library(ggplot2)
library(plotly)

# Bioinformatics libraries
library(biomaRt)
library(org.Hs.eg.db)
```

# 2. Load data
The dataset from this analysis can be downloaded from this FigShare url: https://figshare.com/collections/Integrated_CRISPR/5289226/1

Note that the final batch correction method used to perform downstream analyses was ComBat + QN + PC1, as it was also used in the Pacini et al., 2021 paper for perform downstream analyses. Additionally, we used the CERES method as the preprocessing method. There are additional batch correction and data preprocessing methods that can be used in lieu of these choices if needed.

```{r}
datapath = "D:/Chandrasekaran/Projects/EMT/Data/CRISPR/BatchCorrected/13252640/CERES_SQ_Combat_PC1_All_merge_F.Rds"
data = readRDS(datapath)
```

# 3. Data mapping
The current data matrix is mapped using gene symbols and CCLE identifiers. We should create a data object that can easily map to other gene identifiers and cell line meta data.

## A. Map to CCLE metadata
First, we'll create a map that maps to essential metadata.

### i. Load CCLE metadata
The metadata is available in the CCLE data repository: https://portals.broadinstitute.org/ccle.

**We're interested in the following attributes:**
  * DepMap ID: The cell line identifiers corresponding to the DepMap repository
  * Stripped Cell Line Name: The common cell line name
  * Primary or Metastasis: A categorical describing whether or not the cell line was obtained from the primary tissue site or if it is metastatic
  * Lineage: Tissue of origin
  * Lineage Subtype: Cancer subclassification for tissue of origin
  
```{r}
# Dell
metapath = 'D:/Chandrasekaran/Projects/EMT/Data/Meta/sample_info.csv'

# Read in meta data
metadata = read_csv(metapath)
metadata = metadata[, c("DepMap_ID", 
                        "stripped_cell_line_name", 
                        "primary_or_metastasis",
                        "lineage",
                        "lineage_subtype")]
```

### ii. Filter dataset based on column names
The following code will get the matching metadata `cellmatch` and the matching dataset `match_data`.
```{r}
# Get column names
ids = colnames(data)

# Get match
idx = metadata$DepMap_ID %in% ids
cellmatch = metadata[idx, ]

# Remove non-match in dataset
idx = ids %in% cellmatch$DepMap_ID
match_data = data[, idx]
```

## B. Map to gene identifiers
The following code will match the gene symbols to more robust identifiers: Entrez gene IDs and Ensembl. 

### i. Load up BioMart
First, we'll use the `hsapiens_gene_ensembl` dataset to map the CCLE data.
```{r}
mart = biomaRt::useMart(biomart="ensembl", 
                        dataset="hsapiens_gene_ensembl")
```

### ii. Map gene IDs
Now we can create a map of Ensembl <-> Entrez <-> Symbols. 
```{r}
# Create a map of the probes to Entrez and gene symbols
human_ids = getBM(attributes=c('ensembl_gene_id',
                               'entrezgene_id',
                               'hgnc_symbol'), 
                  values=rownames(match_data),
                  mart=mart)
```

We'll have varying dataset sizes for the metabolic gene sets, depending on how much coverage is available for Ensembl and Entrez. 

### iii. Map to RECON1 genes
First, let's map the data to RECON1, which uses Entrez IDs as the map.

First load the RECON1 genes.
```{r}
recon1_map = "D:/Chandrasekaran/Projects/eGEM/Data/Reconstructions/RECON1/RECON1.xlsx"
recon1_genes = readxl::read_excel(recon1_map, sheet="Genes")
```

Then, we can grab the index that matches between the two datasets
```{r}
# Grab map between Entrez and Symbol
ids = rownames(match_data)
idx = ids %in% recon1_genes$Symbol
```

Finally, grab the match. It will still be mapped to symbols for visualization purposes
```{r}
recon1_match = match_data[idx, ]
dim(recon1_match)
```

### iv. Map to iHUMAN genes
Then, we'll map the data to iHUMAN, which uses Ensembl IDs as the map.

First, load up the iHUMAN map
```{r}
ihuman_map = "D:/Data/Reconstructions/Human1/Human1_SynGO.xlsx"
ihuman_genes = readxl::read_excel(ihuman_map)
```

Then, we can grab the index that matches between the two datasets
```{r}
# Grab map between Entrez and Symbol
ids = rownames(match_data)
idx = ids %in% ihuman_genes$symbol
```

Finally, grab the match. It will still be mapped to symbols for visualization purposes
```{r}
ihuman_match = match_data[idx, ]
dim(recon1_match)
```

## C. Save everything
Let's save everything as an R dataset for other downstream analyses.
```{r}
savepath="D:/Analysis/EMT/data/batched_ceres.Rdata"
save.image(file=savepath)
```


# 4. EDA
Finally, to gain a better understanding of the dataset, let's plot some simple visualizations.

## A. Data distribution using a histogram
The left-tailed distribution show is expected from the CERES Score. The more negative the value, the greater the effect gene deletion has on a given gene. 

```{r, fig.height=10, fig.width=15, dpi=100}
tmp = reshape2::melt(match_data)
ggplot(tmp, aes(x=value, fill=value)) +
  geom_histogram(bins=100) +
  geom_vline(xintercept=0) +
  labs(x="CERES Score", 
       y="Frequency", 
       title="Batch normalized CERES score distribution (matched)")
```

# 5. Query a specific subset of CERES scores
This section queries a specific subset of CERES scores. First, get genes.
```{r}
list_to_query = c("PC", "ASRGL1", "GOT1", "ACSL3", "Vmn2r37", "ACSL5", "ACSL1",
                  "FTCD", "UGP2", "GAPDH", "GAPDHS", "GLS", "GLS2", "HK1",
                  "HKDC1", "GCK", "HK2", "HK3", "HAL", "AQP9", "RPE", 
                  "MDH1", "MDH1B", "UROC1", "HPD", "SLC25A13", "SLC25A12", "GOT2",
                  "SLC16A1", "ACLY", "ACADM", "ACAD8", "SLC25A20", "CYP2E1",
                  "ADK", "ACHY", "AHCYL1", "SLC25A11", "SLC1A6", "SLC1A1",
                  "SLC1A3", "SLC1A2", "SLC1A7", "CAT", "CRAT", "CDO1", "RBKS",
                  "ECHS1", "HADHA", "HADHB", "ENO1", "ENO3", "ENO2", "ELOVL4",
                  "HSD17B4 ", "ACOX1", "ACAA1", "EHHADH", "FASN", "ALDOA",
                  "ALDOB", "ALDOC", "FBP1", "FBP2", "FAH", "GLUL", "LGSN",
                  "HGD", "HIBADH", "HMGCL", "HPRT1", "GSTZ1", "MDH2", "DBT",
                  "BCKDHB", "BCKDHA", "DLD", "PRODH2", "PTDSS1", "SLC16A1", 
                  "PNPLA4", "SLC25A10", "TALDO1", "TKT", "TKTL2", "TKTL1", 
                  "TPI1", "TPIP2", "SLC6A14"
                  )
subset_df = match_data[rownames(match_data) %in% list_to_query, ]
```

Now, get NSCLC cell lines.
```{r}
nsclc_ccl = metadata[metadata$lineage_subtype == "NSCLC", ]
nsclc_ccl = filter(nsclc_ccl, rowSums(is.na(nsclc_ccl)) != ncol(nsclc_ccl))
```

Filter the data for NSCLC cell lines only.
```{r}
subset_df2 = subset_df[, colnames(subset_df) %in% nsclc_ccl$DepMap_ID]
```

Get the median CS across all cell lines.
```{r}
final_nsclc_score = apply(subset_df2[, -1], 1, mean, na.rm=TRUE)
final_nsclc_score = as.data.frame(final_nsclc_score)
rownames(final_nsclc_score) = rownames(subset_df2)
```

# Summary
This data is used for downstream data analyses and validation of the RECON1 metabolic model simulations.