---
title: "Examine top varying metabolic genes"
output: html_notebook
---

```{r}
rm(list=ls())
```

## Summary
This notebook aims to determine whether or not metabolic genes are being over- or underrepresented in the Cook & Vanderhyden dataset

```{r}
library(Seurat)
library(readxl)
```

## 1. Load Cook & Vanderhyden dataset
First, let's load the Cook and Vanderhyden dataset. Assert that we're only dealing with TGF-B induced cells.
```{r}
filepath = 'D:/Chandrasekaran/Projects/EMT/Data/scRNASeq/a549_quantile.rds'
a549 = readRDS(filepath)
Idents(a549) = "Time"
a549 = subset(a549, idents=c("0d", "8h", "1d", "3d", "7d"))
```

Let's also load the RECON1 mapping
```{r}
recon1map = "D:/Chandrasekaran/Projects/eGEM/Data/Reconstructions/RECON1/RECON1.xlsx"
recon1map = readxl::read_excel(recon1map, sheet="Genes")
```

## 2. Examine number of metabolic genes in the top 3000 genes.
We'll use the RECON1 mapping to get the intersection.

```{r}
scaled_data = a549@assays$RNA@scale.data
```

Get the intersection
```{r}
tmp = merge(scaled_data, recon1map, how='inner', by.x=0, by.y='Symbol')
```

There are 318 metabolic genes mapped using this method. What does the distribution look like?
```{r}
library(reshape2)
library(ggplot2)
tmp = melt(tmp)
tmp$color = 'gray'
tmp$color[tmp$value < lims[2]] = 'blue'
tmp$color[tmp$value > lims[4]] = 'red'
tmp$color = as.factor(tmp$color)

lims = quantile(tmp$value)

ggplot(tmp, aes(x=value, fill=color)) +
  geom_histogram(bins=100, aes(fill=color)) +
  xlim(-3, 3) + 
  ggtitle("Scaled Cook & Vanderhyden metabolic genes distribution") +
  xlab("Scaled expression value") + 
  ylab("Frequency") + 
  geom_vline(xintercept=lims[2]) +
  geom_vline(xintercept=lims[4]) +
  scale_color_manual(values = c("blue" = "blue",
                                "gray"="gray",
                                "red"="red")) 
```

Get number of up and down genes across cells.
```{r}
n_up_top3000 = sum(tmp$value > lims[4])
n_dn_top3000 = sum(tmp$value < lims[2])
```

Strange that we get the same number of entries for both up and down.

## 3. Examine number of all metabolic genes in the entire dataset.

First, we preprocess the dataset exactly how they did it in Cook & Vanderhyden.
```{r}
filepath = 'D:/Chandrasekaran/Projects/EMT/Data/scRNASeq/a549_quantile.rds'
a549 = readRDS(filepath)
a549_scale_all = a549
mito.features = grep(pattern="^MT-", 
                      x=rownames(a549_scale_all), 
                      value=TRUE)
percent.mito = Matrix::colSums(GetAssayData(a549_scale_all, 
                                            slot='counts')[mito.features, ]) / Matrix::colSums(GetAssayData(a549_scale_all, slot='counts'))

a549_scale_all[['percent.mito']] = percent.mito
a549_scale_all = subset(a549_scale_all, 
                 subset=nFeature_RNA > 200 & 
                   percent.mito < 0.09)
a549_scale_all = NormalizeData(a549_scale_all)
a549_scale_all = FindVariableFeatures(a549_scale_all, 
                                      selection.method="vst", 
                                      nfeatures=13239)
```

Next, we re-perform the entire analysis we did above.

```{r}
scaled_data = a549_scale_all@assays$RNA@scale.data
```

Get the intersection
```{r}
tmp = merge(scaled_data, recon1map, how='inner', by.x=0, by.y='Symbol')
```

There are 318 metabolic genes mapped using this method. What does the distribution look like?
```{r}
library(reshape2)
library(ggplot2)
tmp = melt(tmp)
tmp$color = 'gray'
tmp$color[tmp$value < lims[2]] = 'blue'
tmp$color[tmp$value > lims[4]] = 'red'
tmp$color = as.factor(tmp$color)

lims = quantile(tmp$value)

ggplot(tmp, aes(x=value, fill=color)) +
  geom_histogram(bins=100, aes(fill=color)) +
  xlim(-3, 3) + 
  ggtitle("Scaled Cook & Vanderhyden metabolic genes distribution") +
  xlab("Scaled expression value") + 
  ylab("Frequency") + 
  geom_vline(xintercept=lims[2]) +
  geom_vline(xintercept=lims[4]) +
  scale_color_manual(values = c("blue" = "blue",
                                "gray"="gray",
                                "red"="red")) 
```

Get number of up and down genes across cells.
```{r}
n_up_top3000 = sum(tmp$value > lims[4])
n_dn_top3000 = sum(tmp$value < lims[2])
```

Based on the results, we don't have to recompute anything. Taking into consideration all genes just results in 1 additional upregulated gene compared to the old analysis.