---
title: "KO Results"
output: html_notebook
---

```{r}
rm(list=ls())
```

## Summary
This notebook visualizes barplots showing the number of knockouts across datasets.
```{r}
library(tidyverse)
```

## 1. Load data
The data contains the knockout growth rates from COBRA simulations for time-course data.
```{r}
path = "D:/Chandrasekaran/Projects/EMT/Data/Supplementary Tables.xlsx"
ko_grate = readxl::read_excel(path, sheet="S. Table 2")

col_names = colnames(ko_grate)
expt_names = col_names[6:22]
ko_scores = ko_grate[, expt_names]
```

## 2. Preprocess data structure
First we need to melt the dataframe.
```{r}
library(reshape2)
ko_score_by_expt = melt(ko_scores)
```

Now we need to create some variables that will be used for the bar plots.
```{r}
library(stringr)
library(janitor)


tmp = data.frame(str_split_fixed(ko_score_by_expt$variable, "_", 2))
ko_score_by_expt = cbind(tmp, ko_score_by_expt)
ko_score_by_expt$counts = ko_score_by_expt$value < 1

freq_tbl = dcast(data=ko_score_by_expt,
               formula = X1 ~ X2,
               fun.aggregate = sum, 
               value.var = "counts")

freq_tbl[freq_tbl == 0] = NA
freq_tbl = melt(freq_tbl)

```
This table contains all
```{r, fig.width = 14, fig.height = 10}
library(ggplot2)
library(extrafont)
library("ggsci")

loadfonts(device = "win")
#dev.new(family = "Arial")
freq_tbl$variable = gsub('48hrs', '48-72hrs', freq_tbl$variable)
freq_tbl$variable = gsub('72hrs', '48-72hrs', freq_tbl$variable)

to_keep = c("1hrs", "24hrs", "48-72hrs")
freq_tbl = freq_tbl[freq_tbl$variable %in% to_keep, ]
```

## 2. Visualize All Lethal Reactions
```{r, fig.width = 14, fig.height = 10}
ggplot(data=freq_tbl, aes(x=variable, y=value, label=value)) +
  geom_bar(aes(fill=variable), stat="identity", na.rm=TRUE) +
  xlab("") +
  ylab("Number of Lethal Reactions (Growth Score < 0)") +
  ggtitle("Lethal reactions predicted from RECON1") + 
  theme_minimal() + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        text=element_text(size=22, family='TT Arial'),
        axis.text.x = element_text(size=22),
        legend.position="none") +
  guides(fill=guide_legend(title="Time")) + 
  scale_fill_npg()
```

```{r, fig.width = 14, fig.height = 10}
ggplot(data=freq_tbl, aes(x=variable, y=value, label=value)) +
  geom_bar(aes(fill=X1), stat="identity", na.rm=TRUE) +
  xlab("") +
  ylab("Number of Lethal Reactions (Growth Score < 0)") +
  ggtitle("Lethal reactions predicted from RECON1") + 
  theme_minimal() + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        text=element_text(size=22, family='TT Arial'),
        axis.text.x = element_text(.size=22)) +
  guides(fill=guide_legend(title="Experiment")) + 
  scale_fill_npg()
```


## 3. Visualize All Lethal Reactions Broken Down by Expt
```{r, fig.width = 14, fig.height = 10}
library(dplyr)
tmp = freq_tbl %>% 
  filter(!grepl('GSE147405', X1))

ggplot(data=tmp, aes(x=X1, y=value, label=value)) +
  geom_bar(aes(fill=variable), stat="identity", na.rm=TRUE) +
  xlab("") +
  ylab("Number of Lethal Reactions (Growth Score < 0)") +
  ggtitle("Lethal reactions predicted from RECON1") + 
  geom_text(size = 3, position = position_stack(vjust = 0.5)) + 
  theme_minimal() + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        text=element_text(size=22, family='TT Arial'),
        axis.text.x = element_text(size=22)) + 
  guides(fill=guide_legend(title="Time")) + 
  scale_fill_npg()
   #scale_fill_manual(values=c('#1E3888','#F5E663','#9C3848'), guide=FALSE)
```
## 4. Visualize All Lethal Reactions Broken Down by bulk v Single-Cell
```{r, fig.width = 10, fig.height = 5}

tmp2 = freq_tbl[!freq_tbl$X1 %in% "GSE147405", ]

ggplot(data=tmp2, aes(x=variable, y=value, label=value)) +
  geom_bar(aes(fill=variable), 
           stat="identity",
           na.rm=TRUE,
           width=1) +
  xlab("") +
  ylab("Number of Lethal Reactions") +
  ggtitle("RECON1 Lethal reactions (Bulk)") + 
  #geom_text(size = 3, position = position_stack(vjust = 0.5)) + 
  theme_minimal() + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        text=element_text(size=25, family='TT Arial'),
        axis.text.x = element_text(size=30),
        axis.text.y = element_text(size=30),
        legend.text=element_text(size=30),
        legend.title=element_text(size=30)) +
  guides(fill=guide_legend(title="Experiments")) + 
  scale_fill_manual(values=c("#CAC4CE", "#725AC1", "#242038")) +
  coord_flip()

```