---
title: "R Notebook"
output: html_notebook
---

## Summary
This notebook processes scRNASeq time course data. Some of the code was adapted from Cook et al., 2020.

```{r}
rm(list=ls())
```

## 1. Load Libraries
```{r, results='hide'}
library(tidyverse)
library(umap)
library(Seurat)
library(writexl)
library(ggpubr)
library(gridExtra)
library(extrafont)
loadfonts(device="win")
```

## 2. Load scRNASeq datasets
Next, we'll load the A549 dataset, which is the lung cancer cell line investigated in GSE147405. Additionally, we'll focus on the $TGF \beta 1$ sample, since this was the way EMT was induced in many bulk experiments.

```{r, include=TRUE}
filepath = 'D:/Chandrasekaran/Projects/EMT/Data/scRNASeq/a549_quantile.rds'
a549 = readRDS(filepath)
```

### A. Visualize Time-Course
This bit of code separates groups by time point and creates individual UMAP plots. This may be useful for later.

```{r, fig.width=24, fig.height=12, results='hide', dpi=300}
colmap = c('0d' = '#97EAD2',
           '8hr'= '#8CC7A1',
           '1d' = '#816E94',
           '3d' = '#74226C',
           '7d' = '#4B2142')
pA = DimPlot(a549, 
        reduction="umap",
        dims=c(1, 2),
        split.by='Time',
        group.by='Time',
        cols=colmap,
        label=FALSE,
        repel=TRUE,
        label.size=4,
        pt.size=5, 
        na.value = "grey50") + 
        labs(title="A549 Time-Course", 
             x="UMAP 1", 
             y="UMAP 2") +
        theme_minimal() +
        theme(plot.title = element_text(size=26),
              legend.text=element_text(size=22),
              axis.title.x=element_text(size=22),
              axis.title.y=element_text(size=22),
              axis.text.x=element_text(size=22),
              axis.text.y=element_text(size=22),
              legend.position=c(0.05, 0.15),
              legend.box="vertical",
              legend.margin=margin(),
              text=element_text(family="sans")) +
        guides(color=guide_legend(override.aes=list(size=2)))

pA
```

This embedding just changes subpopulations to time, which is what I originally wanted.

```{r, fig.width=24, fig.height=12, results='hide', dpi=300}
pB = DimPlot(a549, 
        reduction="umap",
        dims=c(1, 2),
        group.by='Time',
        cols = colmap,
        label=FALSE,
        repel=TRUE,
        label.size=4,
        pt.size=5, 
        na.value = "grey50") + 
        labs(title="A549 Time-Course", 
             x="UMAP 1", 
             y="UMAP 2") +
        theme_minimal() +
        theme(plot.title = element_text(size=26),
              legend.text=element_text(size=22),
              axis.title.x=element_text(size=22),
              axis.title.y=element_text(size=22),
              axis.text.x=element_text(size=22),
              axis.text.y=element_text(size=22),
              legend.position=c(0.05, 0.15),
              legend.box="vertical",
              legend.margin=margin(),
              text=element_text(family="sans")) +
        guides(color=guide_legend(override.aes=list(size=2)))


pB
```

## 3. Overlay metabolic fluxes onto the UMAP embedding
We'll read in the metabolic flux data computed from single-cell Flux Balance Analysis
```{r, results='hide'}
flux_path = 'D:/Chandrasekaran/Projects/EMT/Analysis/old/recon1_flux_quantiles.csv'
flux_data = readr::read_csv(flux_path)

flux_path = 'D:/Chandrasekaran/Projects/EMT/Analysis/old/recon1_rxnko_quantiles.csv'
flux_data = readr::read_csv(flux_path)


col_names = colnames(flux_data)
col_names = stringr::str_replace(col_names, '...[0-9]+', '')
colnames(flux_data) = col_names
#flux_data[, 4:ncol(flux_data)] = abs(flux_data[, 4:ncol(flux_data)])
```

Now let's perform some mapping to the original ids
```{r, results='hide'}
a549_id_map = read_csv("D:/Chandrasekaran/Projects/EMT/Analysis/data/a549_colmap.csv")
```
Now let's do a concatenation. It looks like there is perfect correspondance between the map and the dataset.
```{r}
col_names = data.frame(col_names[4:length(col_names)])
sanity_check = merge(col_names, a549_id_map, by='row.names')
```

Now let's replace the names. This will be used to map to the embedding.
```{r}
colnames(flux_data) = c("BiGG", "Name", "Subsystem", sanity_check$orig_id)
rxn_ids = flux_data[, c(2:3)]
flux_data = flux_data[, -c(2:3)]
flux_data = t(flux_data)

colnames(flux_data) = flux_data[1, ]
flux_data = data.frame(flux_data)
flux_data = flux_data[-1, ] 
```

```{r}
umap_data = data.frame(a549@reductions$umap@cell.embeddings)
flux_data = mutate_all(flux_data, function(x) as.numeric(as.character(x)))

#umap_data$cluster = rownames(a549$Time)
merged_umap = merge(umap_data, flux_data, by.x='row.names', by.y='row.names')
rownames(merged_umap) = merged_umap$Row.names
merged_umap = merged_umap[, -1]
```

Let's intialize some variables we need to declare for the visualization
```{r, results='hide'}
reactions_to_visualize = c("ADSL1", "CYSTS", "PEPCKm",
                           "ACITL", "CYSO", "DHFR")
reaction_names = c("adenylosuccinate lyase", 
                   "cystathionine beta-synthase", 
                   "Phosphoenolpyruvate carboxykinase (GTP)",
                   "ATP-Citrate lyase",
                   "cysteine oxidase",
                   "dihydrofolate reductase")
colors = c("#CC3311", "#CC3311", "#CC3311",
           "#004488", "#004488", "#004488")
```

```{r}
tmp = merged_umap[, reactions_to_visualize]
```

```{r}
p = list()
p[[1]] = ggplot(merged_umap, aes(x=UMAP_1, y=UMAP_2, color=merged_umap[, reactions_to_visualize[1]])) + 
                geom_point(size=3, alpha=0.5) +
                theme_minimal() +
                labs(x="UMAP 1", 
                     y="UMAP 2",
                     title=reaction_names[1], 
                     color="Flux") + 
                scale_color_gradient(low="gray", high=colors[1],
                                     limits = c(0, 1.3)) +
                theme(plot.title = element_text(size=10),
                      legend.text=element_text(size=8),
                      axis.title.x=element_text(size=10),
                      axis.title.y=element_text(size=10),
                      axis.text.x=element_text(size=8),
                      axis.text.y=element_text(size=8),
                      legend.box="vertical",
                      legend.margin=margin())

p[[2]] = ggplot(merged_umap, aes(x=UMAP_1, y=UMAP_2, color=merged_umap[, reactions_to_visualize[2]])) + 
                geom_point(size=3, alpha=0.5) +
                theme_minimal() +
                labs(x="UMAP 1", 
                     y="UMAP 2",
                     title=reaction_names[2], 
                     color="Flux") + 
                scale_color_gradient(low="gray", high=colors[2],
                                     limits = c(0,1.3)) +
                theme(plot.title = element_text(size=10),
                      legend.text=element_text(size=8),
                      axis.title.x=element_text(size=10),
                      axis.title.y=element_text(size=10),
                      axis.text.x=element_text(size=8),
                      axis.text.y=element_text(size=8),
                      legend.box="vertical",
                      legend.margin=margin())

p[[3]] = ggplot(merged_umap, aes(x=UMAP_1, y=UMAP_2, color=merged_umap[, reactions_to_visualize[3]])) + 
                geom_point(size=3, alpha=0.5) +
                theme_minimal() +
                labs(x="UMAP 1", 
                     y="UMAP 2",
                     title=reaction_names[3], 
                     color="Flux") + 
                scale_color_gradient(low="gray", high=colors[3],
                                     limits = c(0,1.3)) +
                theme(plot.title = element_text(size=10),
                      legend.text=element_text(size=8),
                      axis.title.x=element_text(size=10),
                      axis.title.y=element_text(size=10),
                      axis.text.x=element_text(size=8),
                      axis.text.y=element_text(size=8),
                      legend.box="vertical",
                      legend.margin=margin())

p[[4]] = ggplot(merged_umap, aes(x=UMAP_1, y=UMAP_2, color=merged_umap[, reactions_to_visualize[4]])) + 
                geom_point(size=3, alpha=0.5) +
                theme_minimal() +
                labs(x="UMAP 1", 
                     y="UMAP 2",
                     title=reaction_names[4], 
                     color="Flux") + 
                scale_color_gradient(low="gray", high=colors[4],
                                     limits = c(0,1.3)) +
                theme(plot.title = element_text(size=10),
                      legend.text=element_text(size=8),
                      axis.title.x=element_text(size=10),
                      axis.title.y=element_text(size=10),
                      axis.text.x=element_text(size=8),
                      axis.text.y=element_text(size=8),
                      legend.box="vertical",
                      legend.margin=margin())

p[[5]] = ggplot(merged_umap, aes(x=UMAP_1, y=UMAP_2, color=merged_umap[, reactions_to_visualize[5]])) + 
                geom_point(size=3, alpha=0.5) +
                theme_minimal() +
                labs(x="UMAP 1", 
                     y="UMAP 2",
                     title=reaction_names[5], 
                     color="Flux") + 
                scale_color_gradient(low="gray", high=colors[5],
                                     limits = c(0,1.3)) +
                theme(plot.title = element_text(size=10),
                      legend.text=element_text(size=8),
                      axis.title.x=element_text(size=10),
                      axis.title.y=element_text(size=10),
                      axis.text.x=element_text(size=8),
                      axis.text.y=element_text(size=8),
                      legend.box="vertical",
                      legend.margin=margin())

p[[6]] = ggplot(merged_umap, aes(x=UMAP_1, y=UMAP_2, color=merged_umap[, reactions_to_visualize[6]])) + 
                geom_point(size=3, alpha=0.5) +
                theme_minimal() +
                labs(x="UMAP 1", 
                     y="UMAP 2",
                     title=reaction_names[6], 
                     color="Flux") + 
                scale_color_gradient(low="gray", high=colors[6],
                                     limits = c(0,1.3)) +
                theme(plot.title = element_text(size=10),
                      legend.text=element_text(size=8),
                      axis.title.x=element_text(size=10),
                      axis.title.y=element_text(size=10),
                      axis.text.x=element_text(size=8),
                      axis.text.y=element_text(size=8),
                      legend.box="vertical",
                      legend.margin=margin())

```


Create and visualize the final UMAP flux plot object.
```{r, fig.width=24, fig.height=12, results='hide'}
pC = scater::multiplot(p[[1]], p[[6]], p[[2]], 
                       p[[4]], p[[3]], p[[5]], cols=3)
pC
```

First, we need to format the data, so that it matches the DGE Matrix, where rows are reactions and columns are cell IDs.
```{r, results='hide'}
fba = flux_data

# Reformat
fba = t(fba)
fba = as.sparse(fba)
```

Now we need to make sure the expression matrix has the same cells as the fba matrix, and also ensure the `active.ident` and the `Time` datapoints we'll map to.
```{r, results='hide'}
# Get names and indices
intersect_cells = intersect(colnames(a549@assays$RNA@counts), colnames(fba))
cell_idx = which(colnames(a549@assays$RNA@counts) %in% intersect_cells)

# Reformat data
a549_subset = subset(a549, cells=intersect_cells)
fba = fba[, intersect_cells]

a549_subset[["FBA"]] = CreateAssayObject(counts=fba)
saveRDS(a549_subset, file = "D:/Chandrasekaran/Projects/EMT/Data/scRNASeq/a549_quantile.rds")

```

```{r, results='hide'}
preprocess_flux = function(gene_obj, assay){
  # Log normalize data
  gene_obj = NormalizeData(gene_obj, 
                           assay=assay,
                           normalization.method='LogNormalize', 
                           scale.factor=10000)
  
  # Scale data
  gene_obj = ScaleData(gene_obj, 
                       assay=assay)
  return(gene_obj)
}
```

Now let's perform the data preprocessing:
```{r, results='hide'}
#a549_subset = preprocess_flux(a549_subset, assay="FBA")
```

## D. Get Metabolic Markers
We'll get metabolic markers based on subpopulations.
```{r, results='hide'}
a549_subset@active.ident = as.factor(a549_subset@meta.data$Time)
subpop_reactions = FindAllMarkers(a549_subset,
                                  assay="FBA",
                                  only.pos=TRUE,
                                  min.pct=0.25,
                                  logfc.threshold=0.05,
                                  test.use="MAST")
#write.table(subpop_reactions, "clipboard", sep="\t")
```

```{r, results='hide'}
#var_feats = FindVariableFeatures(object=a549_subset)
DoHeatmap(object=a549_subset,
          assay="FBA",
          slot='counts',
          label=FALSE,
          draw.lines=TRUE,
          lines.width=10,
          theme(text=element_text(family='sans', size=16),
                legend.text=element_text(size=16),
                axis.title.x=element_text(size=20),
                axis.title.y=element_text(size=20),
                axis.text.x=element_text(size=16),
                axis.text.y=element_text(size=16),
                legend.title=element_text(size=20)))
```