---
title: "R Notebook"
output: html_notebook
---

## Summary
This notebook processes scRNASeq time course data. Some of the code was adapted from Cook et al., 2020.

```{r}
rm(list=ls())
```

## 1. Load Libraries
```{r, results='hide'}
library(tidyverse)
library(umap)
library(Seurat)
library(writexl)
library(ggpubr)
library(gridExtra)
library(extrafont)
#loadfonts(device="win")
```

## 2. Load scRNASeq datasets
Next, we'll load the A549 dataset, which is the lung cancer cell line investigated in GSE147405. Additionally, we'll focus on the $TGF \beta 1$ sample, since this was the way EMT was induced in many bulk experiments.

```{r, include=TRUE}
filepath = 'D:/Chandrasekaran/Projects/EMT/Data/scRNASeq/a549_quantile.rds'
a549 = readRDS(filepath)
```

### A. Visualize Time-Course
This bit of code separates groups by time point and creates individual UMAP plots. This may be useful for later.

```{r, fig.width=24, fig.height=12, results='hide', dpi=300}
colmap = c('0d' = '#97EAD2',
           '8hr'= '#8CC7A1',
           '1d' = '#816E94',
           '3d' = '#74226C',
           '7d' = '#4B2142')
pA = DimPlot(a549, 
        reduction="umap",
        dims=c(1, 2),
        split.by='Time',
        group.by='Time',
        cols=colmap,
        label=FALSE,
        repel=TRUE,
        label.size=4,
        pt.size=5, 
        na.value = "grey50") + 
        labs(title="A549 Time-Course", 
             x="UMAP 1", 
             y="UMAP 2") +
        theme_minimal() +
        theme(plot.title = element_text(size=26),
              legend.text=element_text(size=22),
              axis.title.x=element_text(size=22),
              axis.title.y=element_text(size=22),
              axis.text.x=element_text(size=22),
              axis.text.y=element_text(size=22),
              legend.position=c(0.05, 0.15),
              legend.box="vertical",
              legend.margin=margin(),
              text=element_text(family="sans")) +
        guides(color=guide_legend(override.aes=list(size=2)))

pA
```

This embedding just changes subpopulations to time, which is what I originally wanted.

```{r, fig.width=24, fig.height=12, results='hide', dpi=300}
pB = DimPlot(a549, 
        reduction="umap",
        dims=c(1, 2),
        group.by='Time',
        cols = colmap,
        label=FALSE,
        repel=TRUE,
        label.size=4,
        pt.size=5, 
        na.value = "grey50") + 
        labs(title="A549 Time-Course", 
             x="UMAP 1", 
             y="UMAP 2") +
        theme_minimal() +
        theme(plot.title = element_text(size=26),
              legend.text=element_text(size=22),
              axis.title.x=element_text(size=22),
              axis.title.y=element_text(size=22),
              axis.text.x=element_text(size=22),
              axis.text.y=element_text(size=22),
              legend.position=c(0.05, 0.15),
              legend.box="vertical",
              legend.margin=margin(),
              text=element_text(family="sans")) +
        guides(color=guide_legend(override.aes=list(size=2)))


pB
```

## 3. Overlay growth rates from KO onto the UMAP embedding
We'll read in the metabolic flux data computed from single-cell Flux Balance Analysis
```{r, results='hide'}
flux_path = 'D:/Chandrasekaran/Projects/EMT/Analysis/old/recon1_flux_quantiles.csv'
flux_data = readr::read_csv(flux_path)

#flux_path = 'D:/Chandrasekaran/Projects/EMT/Analysis/old/recon1_rxnko_quantiles.csv'
#flux_data = readr::read_csv(flux_path)


col_names = colnames(flux_data)
col_names = stringr::str_replace(col_names, '...[0-9]+', '')
colnames(flux_data) = col_names
#flux_data[, 4:ncol(flux_data)] = abs(flux_data[, 4:ncol(flux_data)])
```

Now let's perform some mapping to the original ids
```{r, results='hide'}
a549_id_map = read_csv("D:/Chandrasekaran/Projects/EMT/Analysis/data/a549_colmap.csv")
```

Now let's do a concatenation. It looks like there is perfect correspondance between the map and the dataset.
```{r}
col_names = data.frame(col_names[4:length(col_names)])
sanity_check = merge(col_names, a549_id_map, by='row.names')
```

Now let's replace the names. This will be used to map to the embedding.
```{r}
colnames(flux_data) = c("BiGG", "Name", "Subsystem", sanity_check$orig_id)
rxn_ids = flux_data[, c(2:3)]
flux_data = flux_data[, -c(2:3)]
flux_data = t(flux_data)

colnames(flux_data) = flux_data[1, ]
flux_data = data.frame(flux_data)
flux_data = flux_data[-1, ] 
```

This will perform a zscore transformation
```{r}
flux_data = mutate_all(flux_data, function(x) as.numeric(as.character(x)))
flux_data[is.na(flux_data)] = 0
```

Let's visualize the Z-score distirbution.
```{r}
library(reshape2)
tmp = melt(flux_data)
ggplot(tmp, aes(x=value)) + 
  geom_histogram(bins=100) + 
  scale_y_continuous(trans='log10') +
  ggtitle("RECON1 scRNASeq KO distribution") +
  xlab("Reaction KO Distribution") +
  ylab("Frequency (log-scale)") +
  theme_minimal()
```

```{r}
tmp = scale(flux_data, center=TRUE, scale=TRUE)
```

Get the standard deviations for each column
```{r}
rxn_std = apply(flux_data, 2, sd)
rxn_std_sorted = sort(rxn_std, 
                      decreasing = TRUE, 
                      index.return=TRUE)
```

Let's visualize the Z-score distirbution.
```{r}
tmp = melt(tmp)
ggplot(tmp, aes(x=value)) + 
  geom_histogram(bins=100) + 
  scale_y_continuous(trans='log10') +
  ggtitle("RECON1 scRNASeq KO distribution") +
  xlab("Reaction KO Z-Score Distribution") +
  ylab("Frequency (log-scale)") +
  theme_minimal()
```
This combines the reaction ko data with the UMAP embedding.
```{r}
umap_data = data.frame(a549@reductions$umap@cell.embeddings)

#umap_data$cluster = rownames(a549$Time)
merged_umap = merge(umap_data, flux_data, by.x='row.names', by.y='row.names')
rownames(merged_umap) = merged_umap$Row.names
merged_umap = merged_umap[, -1]
```

This combines the reaction ko z-score data with the UMAP embedding.
```{r}
umap_data = data.frame(a549@reductions$umap@cell.embeddings)

#umap_data$cluster = rownames(a549$Time)
merged_umap = merge(umap_data, tmp, by.x='row.names', by.y='row.names')
rownames(merged_umap) = merged_umap$Row.names
merged_umap = merged_umap[, -1]
```

Let's intialize some variables we need to declare for the visualization
```{r, results='hide'}
reactions_to_visualize = c("ADSL1", "CYSTS", "PEPCKm",
                           "ACITL", "CYSO", "DHFR")
reaction_names = c("adenylosuccinate lyase", 
                   "cystathionine beta-synthase", 
                   "Phosphoenolpyruvate carboxykinase (GTP)",
                   "ATP-Citrate lyase",
                   "cysteine oxidase",
                   "dihydrofolate reductase")

```

```{r}
reactions_to_visualize = c("APAT2rm", "BALAtmr", "MMSAD3m",
                           "PCm", "GAPD", "EX_h_e")
reaction_names = c("3-Aminopropanoate:2-oxoglutarate aminotransferase (m)",
                   "Beta-alanine reversible mitochondrial transport (diffusion)",
                   "methylmalonate-semialdehyde dehydrogenase (malonic semialdehyde), mitochondrial", 
                   "pyruvate carboxylase",
                   "glyceraldehyde-3-phosphate dehydrogenase",
                   "H+ exchange")
```

```{r, results='hide'}
reactions_to_visualize = c("AKGMALtm", "ENO", "ACITL",
                           "CSNATr", "MDHm", "LDH_L")
reaction_names = c("AKG Transporter", 
                   "Enolase", 
                   "ATP Citrate Lyase",
                   "Carnitine o-acetyltransferase",
                   "Malate dehydrogenase")
```

```{r, results='hide'}
reactions_to_visualize = c("PGCD", "PSERT", "TPI",
                           "PCm", "GAPD", "LDH_L")
reaction_names = c("phosphoglycerate dehydrogenase", 
                   "phosphoserine transaminase", 
                   "triose phosphate isomerase",
                   "pyruvate carboxylase",
                   "glyceraldehyde-3-phosphate dehydrogenase",
                   "Lactate dehydrogenase")

```

```{r}
tmp = merged_umap[, reactions_to_visualize]
tmp[is.na(tmp)] = 0
```

```{r}
tmp = scale(tmp, center=TRUE, scale=TRUE)
```

```{r}
p = list()
p[[1]] = ggplot(merged_umap, aes(x=UMAP_1, y=UMAP_2, color=tmp[, reactions_to_visualize[1]])) + 
                geom_point(size=5) +
                theme_minimal() +
                labs(x="UMAP 1", 
                     y="UMAP 2",
                     title=reaction_names[1], 
                     color="Flux") + 
                scale_colour_gradient2(low="red", mid = 'gray', high='black',
                                     limit=c(-2, 1)) +
                theme(plot.title = element_text(size=26),
                            legend.text=element_text(size=22),
                            axis.title.x=element_text(size=22),
                            axis.title.y=element_text(size=22),
                            axis.text.x=element_text(size=22),
                            axis.text.y=element_text(size=22),
                            legend.position=c(0.05, 0.15),
                            legend.box="vertical",
                            legend.margin=margin(),
                            text=element_text(family="sans")) 
                

p[[2]] = ggplot(merged_umap, aes(x=UMAP_1, y=UMAP_2, color=tmp[, reactions_to_visualize[2]])) + 
                geom_point(size=5) +
                theme_minimal() +
                labs(x="UMAP 1", 
                     y="UMAP 2",
                     title=reaction_names[2], 
                     color="Flux") + 
                scale_colour_gradient2(low="red", mid = 'gray', high='black',
                                     limit=c(-2, 1)) +
                theme(plot.title = element_text(size=26),
                            legend.text=element_text(size=22),
                            axis.title.x=element_text(size=22),
                            axis.title.y=element_text(size=22),
                            axis.text.x=element_text(size=22),
                            axis.text.y=element_text(size=22),
                            legend.position=c(0.05, 0.15),
                            legend.box="vertical",
                            legend.margin=margin(),
                            text=element_text(family="sans"))

p[[3]] = ggplot(merged_umap, aes(x=UMAP_1, y=UMAP_2, color=tmp[, reactions_to_visualize[3]])) + 
                geom_point(size=5) +
                theme_minimal() +
                labs(x="UMAP 1", 
                     y="UMAP 2",
                     title=reaction_names[3], 
                     color="Flux") + 
                scale_colour_gradient2(low="red", mid = 'gray', high='black',
                                     limit=c(-2, 1)) +
                theme(plot.title = element_text(size=26),
                            legend.text=element_text(size=22),
                            axis.title.x=element_text(size=22),
                            axis.title.y=element_text(size=22),
                            axis.text.x=element_text(size=22),
                            axis.text.y=element_text(size=22),
                            legend.position=c(0.05, 0.15),
                            legend.box="vertical",
                            legend.margin=margin(),
                            text=element_text(family="sans"))

p[[4]] = ggplot(merged_umap, aes(x=UMAP_1, y=UMAP_2, color=tmp[, reactions_to_visualize[4]])) + 
                geom_point(size=5) +
                theme_minimal() +
                labs(x="UMAP 1", 
                     y="UMAP 2",
                     title=reaction_names[4], 
                     color="Flux") + 
                scale_colour_gradient2(low="red", mid = 'gray', high='black',
                                     limit=c(-2, 1)) +
                theme(plot.title = element_text(size=26),
                            legend.text=element_text(size=22),
                            axis.title.x=element_text(size=22),
                            axis.title.y=element_text(size=22),
                            axis.text.x=element_text(size=22),
                            axis.text.y=element_text(size=22),
                            legend.position=c(0.05, 0.15),
                            legend.box="vertical",
                            legend.margin=margin(),
                            text=element_text(family="sans")) 

p[[5]] = ggplot(merged_umap, aes(x=UMAP_1, y=UMAP_2, color=tmp[, reactions_to_visualize[5]])) + 
                geom_point(size=5) +
                theme_minimal() +
                labs(x="UMAP 1", 
                     y="UMAP 2",
                     title=reaction_names[5], 
                     color="Flux") + 
                scale_colour_gradient2(low="red", mid = 'gray', high='black',
                                     limit=c(-2, 1)) +
                theme(plot.title = element_text(size=26),
                            legend.text=element_text(size=22),
                            axis.title.x=element_text(size=22),
                            axis.title.y=element_text(size=22),
                            axis.text.x=element_text(size=22),
                            axis.text.y=element_text(size=22),
                            legend.position=c(0.05, 0.15),
                            legend.box="vertical",
                            legend.margin=margin(),
                            text=element_text(family="sans")) 

```


Create and visualize the final UMAP flux plot object.
```{r, fig.width=24, fig.height=12, results='hide'}
pC = scater::multiplot(pB, p[[1]], p[[2]], 
                       p[[4]], p[[3]], p[[5]], cols=3)
#pC
```

## 4. Save the data
First, we need to format the data, so that it matches the DGE Matrix, where rows are reactions and columns are cell IDs.
```{r, results='hide'}
fba = flux_data

# Reformat
fba = t(fba)
fba = as.sparse(fba)
```

Now we need to make sure the expression matrix has the same cells as the fba matrix, and also ensure the `active.ident` and the `Time` datapoints we'll map to.
```{r, results='hide'}
# Get names and indices
intersect_cells = intersect(colnames(a549@assays$RNA@counts), colnames(fba))
cell_idx = which(colnames(a549@assays$RNA@counts) %in% intersect_cells)

# Reformat data
a549_subset = subset(a549, cells=intersect_cells)
fba = fba[, intersect_cells]

a549_subset[["FBA"]] = CreateAssayObject(counts=fba)
saveRDS(a549_subset, file = "D:/Chandrasekaran/Projects/EMT/Data/scRNASeq/a549_quantile.rds")

```

Let's create a preprocessing pipeline for downstream analyses
```{r, results='hide'}
preprocess_flux = function(gene_obj, assay){
  # Log normalize data
  gene_obj = NormalizeData(gene_obj, 
                           assay=assay,
                           normalization.method='LogNormalize', 
                           scale.factor=10000)
  
  # Scale data
  gene_obj = ScaleData(gene_obj, 
                       assay=assay)
  return(gene_obj)
}
```

Now let's perform the data preprocessing
```{r, results='hide'}
#a549_subset = preprocess_flux(a549_subset, assay="FBA")
```

## 5. Get Metabolic Markers
We'll get metabolic markers based on subpopulations. This module did not yield any result
```{r, results='hide'}
a549_subset@active.ident = as.factor(a549_subset@meta.data$Time)
subpop_reactions = FindAllMarkers(a549_subset,
                                  assay="FBA",
                                  only.pos=TRUE,
                                  min.pct=0.25,
                                  logfc.threshold=0.05,
                                  test.use="MAST")
#write.table(subpop_reactions, "clipboard", sep="\t")
```

```{r, results='hide'}
#var_feats = FindVariableFeatures(object=a549_subset)
DoHeatmap(object=a549_subset,
          assay="FBA",
          slot='counts',
          label=FALSE,
          draw.lines=TRUE,
          lines.width=10,
          theme(text=element_text(family='sans', size=16),
                legend.text=element_text(size=16),
                axis.title.x=element_text(size=20),
                axis.title.y=element_text(size=20),
                axis.text.x=element_text(size=16),
                axis.text.y=element_text(size=16),
                legend.title=element_text(size=20)))
```

## 6. Overlay metabolic fluxes 
Now let's repeat for metabolic fluxes.
```{r}
flux_path = 'D:/Chandrasekaran/Projects/EMT/Analysis/old/recon1_flux_quantiles.csv'
flux_data = readr::read_csv(flux_path)


col_names = colnames(flux_data)
col_names = stringr::str_replace(col_names, '...[0-9]+', '')
colnames(flux_data) = col_names
#flux_data[, 4:ncol(flux_data)] = abs(flux_data[, 4:ncol(flux_data)])
```

Now let's perform some mapping to the original ids
```{r, results='hide'}
a549_id_map = read_csv("D:/Chandrasekaran/Projects/EMT/Analysis/data/a549_colmap.csv")
```

Now let's do a concatenation. It looks like there is perfect correspondance between the map and the dataset.
```{r}
col_names = data.frame(col_names[4:length(col_names)])
sanity_check = merge(col_names, a549_id_map, by='row.names')
```

Now let's replace the names. This will be used to map to the embedding.
```{r}
colnames(flux_data) = c("BiGG", "Name", "Subsystem", sanity_check$orig_id)
rxn_ids = flux_data[, c(1:3)]
glycolysis_ids = rxn_ids[rxn_ids$Subsystem=='Glycolysis/Gluconeogenesis', 1]
glycolysis_ids = na.omit(glycolysis_ids)
glycolysis_data = flux_data[flux_data$Subsystem=='Glycolysis/Gluconeogenesis', ]
glycolysis_data = glycolysis_data[, -c(2:3)]
glycolysis_data = t(glycolysis_data)

colnames(glycolysis_data) = glycolysis_data[1, ]
glycolysis_data = data.frame(glycolysis_data)
glycolysis_data = glycolysis_data[-1, ] 
```

```{r}
glycolysis_data = glycolysis_data[ , apply(glycolysis_data, 2, function(x) !all(is.na(x)))]
```

```{r}
umap_data = data.frame(a549@reductions$umap@cell.embeddings)
glycolysis_data = mutate_all(glycolysis_data, function(x) as.numeric(as.character(x)))
glycolysis_data = normalize(glycolysis_data)

#umap_data$cluster = rownames(a549$Time)
merged_umap = merge(umap_data, glycolysis_data, by.x='row.names', by.y='row.names')
rownames(merged_umap) = merged_umap$Row.names
merged_umap = merged_umap[, -1]
```

```{r}
normalize <- function(x, na.rm = TRUE) {
    return((x- min(x)) /(max(x)-min(x)))
}
```

```{r}
col = merged_umap[as.character(glycolysis_ids[5,])]
p = list()
p[[1]] = ggplot(merged_umap, aes(x=UMAP_1, y=UMAP_2, color=ENO)) + 
                geom_point(size=5, alpha=0.75) +
                theme_minimal() +
                labs(x="UMAP 1", 
                     y="UMAP 2",
                     title="ENO", 
                     color="Flux") + 
                scale_colour_gradient2(high="#D00000", mid='#F48C06', low='gray', midpoint=0.5, limit=c(0,1)) +
                theme(plot.title = element_text(size=26),
                            legend.text=element_text(size=22),
                            axis.title.x=element_text(size=22),
                            axis.title.y=element_text(size=22),
                            axis.text.x=element_text(size=22),
                            axis.text.y=element_text(size=22),
                            legend.position=c(0.05, 0.15),
                            legend.box="vertical",
                            legend.margin=margin(),
                            text=element_text(family="sans")) 

p[[2]] = ggplot(merged_umap, aes(x=UMAP_1, y=UMAP_2, color=HEX1)) + 
                geom_point(size=5, alpha=0.75) +
                theme_minimal() +
                labs(x="UMAP 1", 
                     y="UMAP 2",
                     title="HK", 
                     color="Flux") + 
                scale_colour_gradient2(high="#D00000", mid='#F48C06', low='gray', midpoint=0.5, limit=c(0,1)) +
                theme(plot.title = element_text(size=26),
                            legend.text=element_text(size=22),
                            axis.title.x=element_text(size=22),
                            axis.title.y=element_text(size=22),
                            axis.text.x=element_text(size=22),
                            axis.text.y=element_text(size=22),
                            legend.position=c(0.05, 0.15),
                            legend.box="vertical",
                            legend.margin=margin(),
                            text=element_text(family="sans")) 

p[[3]] = ggplot(merged_umap, aes(x=UMAP_1, y=UMAP_2, color=GAPD)) + 
                geom_point(size=5, alpha=0.75) +
                theme_minimal() +
                labs(x="UMAP 1", 
                     y="UMAP 2",
                     title="GAPD", 
                     color="Flux") + 
                scale_colour_gradient2(high="#D00000", mid='#F48C06', low='gray', midpoint=0.5, limit=c(0,1)) +
                theme(plot.title = element_text(size=26),
                            legend.text=element_text(size=22),
                            axis.title.x=element_text(size=22),
                            axis.title.y=element_text(size=22),
                            axis.text.x=element_text(size=22),
                            axis.text.y=element_text(size=22),
                            legend.position=c(0.05, 0.15),
                            legend.box="vertical",
                            legend.margin=margin(),
                            text=element_text(family="sans")) 

p[[4]] = ggplot(merged_umap, aes(x=UMAP_1, y=UMAP_2, color=FBA)) + 
                geom_point(size=5, alpha=0.75) +
                theme_minimal() +
                labs(x="UMAP 1", 
                     y="UMAP 2",
                     title="FBA", 
                     color="Flux") + 
                scale_colour_gradient2(high="#D00000", mid='#F48C06', low='gray', midpoint=0.5, limit=c(0,1)) +
                theme(plot.title = element_text(size=26),
                            legend.text=element_text(size=22),
                            axis.title.x=element_text(size=22),
                            axis.title.y=element_text(size=22),
                            axis.text.x=element_text(size=22),
                            axis.text.y=element_text(size=22),
                            legend.position=c(0.05, 0.15),
                            legend.box="vertical",
                            legend.margin=margin(),
                            text=element_text(family="sans")) 

p[[5]] = ggplot(merged_umap, aes(x=UMAP_1, y=UMAP_2, color=TPI)) + 
                geom_point(size=5, alpha=0.75) +
                theme_minimal() +
                labs(x="UMAP 1", 
                     y="UMAP 2",
                     title="TPI", 
                     color="Flux") + 
                scale_colour_gradient2(high="#D00000", mid='#F48C06', low='gray', midpoint=0.5, limit=c(0,1)) +
                theme(plot.title = element_text(size=26),
                            legend.text=element_text(size=22),
                            axis.title.x=element_text(size=22),
                            axis.title.y=element_text(size=22),
                            axis.text.x=element_text(size=22),
                            axis.text.y=element_text(size=22),
                            legend.position=c(0.05, 0.15),
                            legend.box="vertical",
                            legend.margin=margin(),
                            text=element_text(family="sans")) 

```

```{r, fig.width=24, fig.height=12, results='hide'}
pD = scater::multiplot(pB, p[[1]], p[[2]], 
                       p[[4]], p[[3]], p[[5]], cols=3)
```