---
title: "Expression Profiles"
output: html_notebook
---

## Summary
This notebook processes scRNASeq time course data. Some of the code was adapted from Cook et al., 2020.

```{r}
rm(list=ls())
```

## 1. Load Libraries
```{r, results='hide'}
#install.packages('tidyverse')
#install.packages("Seurat")
library(tidyverse)
library(Seurat)
library(writexl)
library(ggpubr)
library(gridExtra)
library(extrafont)
loadfonts(device="win")
```

## 2. Load scRNASeq datasets
Next, we'll load the A549 dataset, which is the lung cancer cell line investigated in GSE147405. Additionally, we'll focus on the $TGF \beta 1$ sample, since this was the way EMT was induced in many bulk experiments.

```{r, include=TRUE}
filepath = 'D:/Chandrasekaran/Projects/EMT/Data/scRNASeq/a549_quantile.rds'
a549 = readRDS(filepath)

DefaultAssay(object = a549) <- "RNA"
```

now let's visualize the time-course
```{r}
colmap = c('0d' = '#97EAD2',
           '8hr'= '#8CC7A1',
           '1d' = '#816E94',
           '3d' = '#74226C',
           '7d' = '#4B2142')
pB = DimPlot(a549, 
        reduction="umap",
        dims=c(1, 2),
        group.by='Time',
        cols = colmap,
        label=FALSE,
        repel=TRUE,
        label.size=4,
        pt.size=5, 
        na.value = "grey50") + 
        labs(title="A549 Time-Course", 
             x="UMAP 1", 
             y="UMAP 2") +
        theme_minimal() +
        theme(plot.title = element_text(size=26),
              legend.text=element_text(size=22),
              axis.title.x=element_text(size=22),
              axis.title.y=element_text(size=22),
              axis.text.x=element_text(size=22),
              axis.text.y=element_text(size=22),
              legend.position=c(0.05, 0.15),
              legend.box="vertical",
              legend.margin=margin(),
              text=element_text(family="sans")) +
        guides(color=guide_legend(override.aes=list(size=2)))


pB
```

## 3. Visualize Ecad and Vim
```{r}
p1 = FeaturePlot(a549, features=c("CDH1", "VIM"), combine=TRUE)
p1
#fix.sc = scale_color_gradientn( colours = c("gray", "red"),  limits = c(1, 3))
#p2 = lapply(p1, function (x) x + fix.sc)
#CombinePlots(p2)

```
```{r}
p = list()
umap_embed = merge(as.data.frame(a549@reductions$umap@cell.embeddings), as.data.frame(a549@assays$RNA@scale.data))
tmp = as.data.frame(a549@assays$RNA@scale.data)
p[[1]] = ggplot(umap_embed, aes(x=UMAP_1, y=UMAP_2, color=merged_umap[, reactions_to_visualize[1]])) + 
                geom_point(size=3, alpha=0.5) +
                theme_minimal() +
                labs(x="UMAP 1", 
                     y="UMAP 2",
                     title=reaction_names[1], 
                     color="Flux") + 
                scale_color_gradient(low="gray", high=colors[1],
                                     limits = c(0, 1.3)) +
                theme(plot.title = element_text(size=10),
                      legend.text=element_text(size=8),
                      axis.title.x=element_text(size=10),
                      axis.title.y=element_text(size=10),
                      axis.text.x=element_text(size=8),
                      axis.text.y=element_text(size=8),
                      legend.box="vertical",
                      legend.margin=margin())

p[[2]] = ggplot(merged_umap, aes(x=UMAP_1, y=UMAP_2, color=merged_umap[, reactions_to_visualize[2]])) + 
                geom_point(size=3, alpha=0.5) +
                theme_minimal() +
                labs(x="UMAP 1", 
                     y="UMAP 2",
                     title=reaction_names[2], 
                     color="Flux") + 
                scale_color_gradient(low="gray", high=colors[2],
                                     limits = c(0,1.3)) +
                theme(plot.title = element_text(size=10),
                      legend.text=element_text(size=8),
                      axis.title.x=element_text(size=10),
                      axis.title.y=element_text(size=10),
                      axis.text.x=element_text(size=8),
                      axis.text.y=element_text(size=8),
                      legend.box="vertical",
                      legend.margin=margin())

p[[3]] = ggplot(merged_umap, aes(x=UMAP_1, y=UMAP_2, color=merged_umap[, reactions_to_visualize[3]])) + 
                geom_point(size=3, alpha=0.5) +
                theme_minimal() +
                labs(x="UMAP 1", 
                     y="UMAP 2",
                     title=reaction_names[3], 
                     color="Flux") + 
                scale_color_gradient(low="gray", high=colors[3],
                                     limits = c(0,1.3)) +
                theme(plot.title = element_text(size=10),
                      legend.text=element_text(size=8),
                      axis.title.x=element_text(size=10),
                      axis.title.y=element_text(size=10),
                      axis.text.x=element_text(size=8),
                      axis.text.y=element_text(size=8),
                      legend.box="vertical",
                      legend.margin=margin())

pC = scater::multiplot(p[[1]], p[[2]], p[[3]] cols=3)
pC
```

## 4. Visualize Glycolytic genes

```{r}
#BiocManager::install("clusterProfiler")
library(clusterProfiler)
```

First, we need to get gene symbols.
 

```{r}
glycolytic_genes = read.csv('C:/Users/Scott/Desktop/glycolysis.txt', sep='\n')
glycolytic_genes = as.character(glycolytic_genes$KEGG_GLYCOLYSIS_GLUCONEOGENESIS)
p1 = FeaturePlot(a549, features=glycolytic_genes, combine=TRUE)
p1
#fix.sc = scale_color_gradientn( colours = c("gray", "red"),  limits = c(1, 3))
#p2 = lapply(p1, function (x) x + fix.sc)
#CombinePlots(p2)

```

```{r}
p1 = FeaturePlot(a549, features=c('ENO1', 'ENO2', 'ENO3'), pt.size=5, combine=TRUE)
p1
```


## 5. OXPHOS zgenes
First, we need to get gene 

```{r}
library(stringr)
nadh_dehydrog = str_subset(oxphos_genes, "NDUFA")
p1 = FeaturePlot(a549, features=nadh_dehydrog, combine=TRUE)
p1
```

```{r}
atp_synth = str_subset(oxphos_genes, "ATP")
p1 = FeaturePlot(a549, features=atp_synth, combine=TRUE)
p1
```


```{r}
oxphos_genes = read.csv('C:/Users/Scott/Desktop/oxphos.txt', sep='\n')
oxphos_genes = as.character(oxphos_genes$KEGG_OXIDATIVE_PHOSPHORYLATION)

p1 = FeaturePlot(a549, features=oxphos_genes, combine=TRUE)
p1
#fix.sc = scale_color_gradientn( colours = c("gray", "red"),  limits = c(1, 3))
#p2 = lapply(p1, function (x) x + fix.sc)
#CombinePlots(p2)

```

## 6. Cell Cycle Genes
First, we need to get gene symbols.
```{r}
cellcycle_genes = read.csv('C:/Users/Scott/Desktop/cellcycle.txt', sep='\n')
cellcycle_genes = as.character(cellcycle_genes$KEGG_CELL_CYCLE)

p1 = FeaturePlot(a549, features=cellcycle_genes, combine=TRUE)
p1
#fix.sc = scale_color_gradientn( colours = c("gray", "red"),  limits = c(1, 3))
#p2 = lapply(p1, function (x) x + fix.sc)
#CombinePlots(p2)

```