---
title: "Time-course aggregation for CBM"
output: html_notebook
---

# Summary
This notebook processes scRNASeq time course data. Some of the code was adapted from Cook et al., 2020.

# Download all dependencies

First, we'll download several dependencies for data processing and data visualization.

```{r, include=FALSE}
pkgLst = c("Seurat", "dplyr", "viridis", 
           "matrixStats", "UpSetR", 
           "mgcv", "tidyverse", "reshape2", "ggplot2")
install.packages(pkgLst)

library(mgcv)
library(splines)
library(Seurat)
library(dplyr)
library(tidyr)
library(viridis)
library(matrixStats)
library(UpSetR)
library(reshape2)
library(ggplot2)
```

# Load scRNASeq datasets

Next, we'll load the A549 dataset, which is the lung cancer cell line investigated in GSE147405. Additionally, we'll focus on the $TGF \beta 1$ sample, since this was the way EMT was induced in many bulk experiments.

```{r, include=TRUE}
a549 = readRDS("~/Data/scRNASeq/GSE147405/A549_TGFB1.rds")
```

# Differential expression across all time points

This code block gets differentially expressed genes from the preprocessed Seurat object.

## Bulk scRNASeq across time points

Cook et al performed a GAM with loess on the time covariate for differential expression. Below is the modified `diffExp` function that takes in the Seurat object to get the result.

```{r, include=TRUE}
diffExp = function(seurat_object){
  cells_keep = colnames(seurat_object)[-which(seurat_object$Time %in%                                        c("8h_rm", "1d_rm", "3d_rm"))]
  subset_seurat = subset(seurat_object, cells=cells_keep)
  subset_seurat$Timepoint = as.numeric(gsub("[^0-9.-]", "",                                                       subset_seurat$Time))
  subset_seurat$Timepoint[subset_seurat$Timepoint==8] = 0.33 
  
  exp = as.matrix(GetAssayData(subset_seurat, slot="data")) 
  
  # Fit natural splines to the dataset
  gam = apply(exp, 1, function(x){
    data = data.frame(exp=x, timepoint=subset_seurat$Timepoint)
    temp = gam(exp ~ ns(timepoint, df=3), data=data)
    res = data.frame(Intercept=temp$coefficients[1],
                      Spline_1=temp$coefficients[2],
                      Spline_2=temp$coefficients[3],
                      Spline_3=temp$coefficients[4],
                      pval=summary(temp)$p.table[,4])
  })
  
  # Compute adjusted P-value using FDR
  results <- do.call("rbind", gam)
  results$qval <- p.adjust(results$pval, method="fdr")
  results$Gene <- rownames(results)
  return(results)
}
```

Compute DGE for A549 TGF treatment and save it to a csv file.
```{r, include=TRUE}
# Perform DEA
a549_tgfb1_dge = diffExp(a549)

# Write to file
filepath = "~/Data/Expression/GSE147405/dge_a549_tgfb1.csv"
write.csv(a549_tgfb1_dge, file=filepath, quote=F)
```

Let's now get a list of significant genes with a FDR-corrected p-value less than or equal to 0.01. This will be saved to the file a549_tgfb1_sig.csv.

```{r, include=TRUE}
a549_tgfb1_sig = filter(a549_tgfb1_dge, qval <= 0.01)

# Write to file
filepath = "~/Data/Expression/GSE147405/a549_tgfb1_sig.csv"
write.csv(a549_tgfb1_sig, file=filepath, quote=F)
```
