---
title: "CRISPR-Cas9 Validation"
author: "Scott Campit"
output: html_notebook
---

# Summary
This notebook does the following analysis:
  1. Separate out lung cancer cell lines
  2. Compute differentially expressed CS
  3. Compare it to current iteration of metabolic reactions and genes predicted by COBRA

# 1. Load libraries
```{r}
library(tidyr)
library(readr)
library(stringr)
library(ggplot2)
library(reticulate)
library(limma)
library(Glimma)
library(edgeR)
library(tidyverse)
library(dplyr)
library(reshape2)
library(readxl)
library(ggVennDiagram)
library(rrcov)
```

# 2. Read in data
```{r}
rm(list=ls())
robpca_data = "D:/Analysis/EMT/ceres_robpca.RData"
load(robpca_data)
```

## Grab CCLE meta data
First, we'll get relevant meta data
```{r}
# Linux
#metapath = "~/Data/CRISPR/sample_info.csv"

# Dell
metapath = 'D:/Data/Mappings/CCLE/sample_info.csv'

# Read in meta data
meta_data = read_csv(metapath)
meta_data = meta_data[, c("DepMap_ID", 
                        "stripped_cell_line_name", 
                        "primary_or_metastasis",
                        "lineage",
                        "lineage_subtype")]
```

# 4. Compute P-values from CERES Score

## Global p-value
Next, we'll compute the p-values using the data itself. Most of the data lies close to 0 (no activity), while the significant data points lie on the left side of the data distribution. Thus, we'll compute the left-tailed P-value.
```{r}
mu = mean(ceres, na.rm=TRUE)
stdev = sd(as.matrix(ceres), na.rm=TRUE)
global_pvals = pnorm(as.numeric(ceres), 
                     mean=mu, 
                     sd=stdev,
                     lower.tail=TRUE)
pvals = global_pvals
pvals = data.matrix(pvals)
```

Because we are testing several hypotheses, we should control using the FDR.
```{r}
# Perform FDR Adjustment
adjusted_pval = p.adjust(pvals, method="fdr")
pvals = adjusted_pval
rm(list=c("adjusted_pval", "global_pvals"))
```

## Plot the data distribution

Preprocess data
```{r}
tmp = data.frame(pvals)
tmp$col = as.factor(tmp$pvals < 0.05)
mlt_ceres = reshape2::melt(ceres)
tmp$ceres = mlt_ceres$value
```

Now let's plot the data distribution.
```{r}
# Plot p-value distribution
ggplot(data.frame(tmp), aes(x=pvals, fill=col)) +  
  geom_histogram(bins=100) +
  scale_fill_manual(values = c("TRUE"="red", "FALSE"="gray")) +
  labs(title="CERES FDR distribution (all)", 
       x="FDR Adjusted P-values (left-sided)", 
       y="Frequency") +
  theme(legend.position="none")

# Plot CERES distribution
ggplot(tmp, aes(x=ceres, fill=col)) +  
  geom_histogram(bins=100) +
  scale_fill_manual(values = c("TRUE"="red", "FALSE"="gray")) +
  labs(title="CERES distribution (all)", 
       x="CERES", 
       y="Frequency") +
  theme(legend.position="none")
```

# 5. Map the CERES scores back to CCLE
The data is currently mapped to a) DepMap identifiers, and b) a hybrid of gene symbols and Entrez IDs.

What we need to do is map it to the following:

  * Cancer Cell Line
  * Cancer Cell Line Tissue Lineage
  * Entrez ID

## Filter the data to get only NSCLC
Now we need to select specific cell lines that we're interested in from the NSCLC subtype.
```{r}
# Lung data needs to be NSCLC and we need to know whether it's from the primary site or if it is metastatic
lung_meta_data = meta_data[meta_data["lineage"]=="lung",]
lung_meta_data = lung_meta_data[lung_meta_data["lineage_subtype"]=="NSCLC",]
lung_meta_data = lung_meta_data[!is.na(lung_meta_data$primary_or_metastasis),]
``` 

## Separate entrez and gene symbol identifiers
We should clean up the column names for the Project Achilles and store as both symbol and entrez ID separately.
```{r}
# Save gene symbols
orig_colnames = colnames(ceres)
all_symbol = as.character(sapply(strsplit(orig_colnames, " "), '[', 1))
#all_symbol = all_symbol[-1]

# Save Entrez ID 
all_entrez = as.character(sapply(strsplit(orig_colnames, " "), '[', -1))
all_entrez = str_remove(all_entrez, "[( )]")
all_entrez = str_remove(all_entrez, "[( )]")
all_entrez[1] = all_symbol[1]
```

Combine all the identifiers into a single dataframe.
```{r}
# All identifiers
all_ids = data.frame(symbol=all_symbol, entrez=all_entrez)
all_ids = all_ids[-1, ]
```

### Map dataset to Entrez IDs
Map everything to Entrez IDs
```{r}
# Create dataset with gene symbols as ID
data_entrez = ceres
colnames(data_entrez) = all_entrez

# Create pvalue dataframe
pvals = as.matrix(pvals)
data_dims = dim(ceres)
pvalue_entrez = matrix(pvals, 
                   nrow=data_dims[1], 
                   ncol=data_dims[2])
pvalue_entrez = data.frame(pvalue_entrez)
colnames(pvalue_entrez) = all_entrez[2:length(pvalue_entrez)] 
```

### Map dataset to Gene Symbols
Map everything to Gene Symbols
```{r}
# Create dataset with gene symbols as ID
data_symbol = ceres
colnames(data_symbol) = all_symbol

# Create pvalue dataframe
pvals = as.matrix(pvals)
data_dims = dim(ceres)
pvalue_symbol = matrix(pvals, 
                   nrow=data_dims[1], 
                   ncol=data_dims[2])
pvalue_symbol = data.frame(pvalue_symbol)
colnames(pvalue_symbol) = all_symbol[2:length(pvalue_symbol)] 
pvalue_symbol$DepMap_ID = ids
```

## Filter metabolic genes only
Map it back to the human metabolic reconstruction.
```{r}
recon1_map = "D:/Data/Mappings/Reconstructions/RECON1/RECON1.xlsx"
recon1_genes = readxl::read_excel(recon1_map, sheet="Genes")
recon1_match = merge(all_ids, recon1_genes, by.x="entrez", by.y="Entrez")

data_entrez = data_entrez[, unique(recon1_match$entrez)]
data_entrez$DepMap_ID = ids

pvalue_entrez = pvalue_entrez[, unique(recon1_match$entrez)]
pvalue_entrez$DepMap_ID = ids
```

# 6. Merge metadata to metabolic p-values and CERES scores
Finally, we can the NSCLC data back to RECON1 genes.
```{r}
# Entrez IDs
merged_lung = merge(lung_meta_data, data_entrez, by='DepMap_ID')
merged_pval = merge(lung_meta_data, pvalue_entrez, by='DepMap_ID')

# Gene Symbols
merged_lung = merge(lung_meta_data, data_symbol, by='DepMap_ID')
merged_pval = merge(lung_meta_data, pvalue_symbol, by='DepMap_ID')

```
## Visualize the NSCLC CRISPR-Cas9 metabolism data mapped to RECON1
Let's construct the data object used for the visualization.
```{r}
tmp = merged_pval[, 6:ncol(merged_pval)]
tmp = reshape2::melt(data.frame(tmp))
tmp$col = as.factor(tmp$value < 0.05)
lung_ceres = reshape2::melt(merged_lung[, 6:ncol(merged_lung)])
tmp$ceres = lung_ceres$value
```

Now actually plot the p-values and CERES scores.
```{r}
# Plot p-value distribution
ggplot(data.frame(tmp), aes(x=value, fill=col)) +  
  geom_histogram(bins=100) +
  scale_fill_manual(values = c("TRUE"="red", "FALSE"="gray")) +
  labs(title="CERES NSCLC p-value distribution", 
       x="P-values (left-sided)", 
       y="Frequency") +
  theme(legend.position="none")

# Plot CERES distribution
ggplot(tmp, aes(x=ceres, fill=col)) +  
  geom_histogram(bins=100) +
  scale_fill_manual(values = c("TRUE"="red", "FALSE"="gray")) +
  labs(title="CERES NSCLC distribution", 
       x="CERES", 
       y="Frequency") +
  theme(legend.position="none")
```

# 7. Get differentially expressed CERES score by comparing metastatic / primary site using limma-voom

## Data preprocessing
Format the dataset
```{r}
sample_info = merged_lung[, 1:5]
data = merged_lung[, -c(1:5)]
rownames(data) = sample_info[, 2]
data = t(data)
sample_info = sample_info %>% dplyr::select(2:3)
```

Now we'll floor positive CERES Scores to 0, and take the absolute value for the CERES Scores
```{r}
data[data > 0] = 0
data = abs(data)
data[is.na(data)] = 0
```

## Preparing limma-voom data 
Now let's preprocess the data
```{r}
# Create a DGE List data object
data = cpm(data)
d0 = DGEList(data)
d0 = calcNormFactors(d0)

# Filter out bad data points
cutoff = 1
drop = which(apply(cpm(d0), 1, max) < cutoff)
d = d0[-drop,] 
dim(d)
```

We need to set up the groups we want to compare and the design matrix.
```{r}
group = as.factor(sample_info$primary_or_metastasis)
design = model.matrix(~0+group)
```

## Run limma-voom
Now let's run *limma-voom*
```{r}
v = voom(d, design, plot=TRUE)
fit = lmFit(v, design, method='robust')
result = eBayes(fit, robust=TRUE)
```

Finally, let's get the top 20 results
```{r}
top.table = topTable(result)
head(top.table, 20)
write.table(top.table, "clipboard", sep="\t", col.names=TRUE)
```

# 8. Get gene list that are universally lethal across all cell lines
This gene list was determined from Project Score. Specifically, I defined universally lethal genes that are lethal in more than 50% of the cancer cell lines out of > 300. 
```{r}
project_score_file = "D:/Data/CRISPR/Score/41586_2019_1103_MOESM4_ESM/Supplementary Table 3 - Core-Fitness and Context-Specificity Summary_v2.xlsx"
sheet_name = "Filter Genes"

project_score = read_excel(project_score_file, sheet_name)
genes_to_filter = project_score[project_score$`Dependency Score` > 50, ]
```

# 9. Get differentially expressed genes for A549 from CERES score
First, we'll get the row corresponding to the cancer cell line A549.
```{r}
a549_ceres = merged_lung[merged_lung$stripped_cell_line_name=="A549",]
a549_ceres[2, ] = merged_pval[merged_pval$stripped_cell_line_name=="A549",]
a549_ceres = a549_ceres[, 6:ncol(a549_ceres)]
a549_ceres = t(a549_ceres)
a549_ceres = data.frame(a549_ceres)
colnames(a549_ceres) = c("CERES", "PValue")
```

Now, let's grab the metabolic genes that are in RECON1
```{r}
xname = data.frame(entrez=rownames(a549_ceres))
tmp = merge(xname, recon1_match, by='entrez')
tmp[, 3:4] = NULL
tmp = unique(tmp)
a549_ceres$Symbol = tmp$symbol
```

## Plot Venn Diagram
```{r}
gene_overlap = list(
  A549_specific = sig_a549_ceres$Symbol,
  Core_fitness = genes_to_filter$`HGNC Symbol`
)
ggVennDiagram(gene_overlap)
```

Finally, we'll grab the genes that have a left-tailed p-value less than 0.05
```{r}
sig_a549_ceres = a549_ceres[a549_ceres$PValue < 0.05, ]
genes_to_filter = genes_to_filter$`Entrez ID`
genes_to_filter = data.frame(genes_to_filter)
ubiq_genes = merge(genes_to_filter, 
                   sig_a549_ceres, 
                   by.x='genes_to_filter', 
                   by.y="row.names")
unique_genes = sig_a549_ceres[!sig_a549_ceres$Symbol %in% ubiq_genes$Symbol, ]
write.table(sig_a549_ceres, "clipboard", sep="\t", col.names=TRUE)
```



# 10. Get differentially expressed genes across all NSCLC cell lines
```{r}
tmp = merged_pval[, -c(1:5)]
tmp2 = merged_lung[, -c(1:5)]
```