---
title: "Processing GSE147405 scRNASeq data for integration with COBRA"
author: "Scott Campit"
output: html_notebook
---

# Summary
This notebook processes scRNASeq time course data. Some of the code was adapted from Cook et al., 2020.

# 1. Load Libraries

First, we'll download several dependencies for data processing and data visualization.

```{r, include=FALSE}
library(Seurat)
library(dplyr)
library(tidyr)
library(reshape2)
library(ggplot2)
library(readxl)
library(biomaRt)
```

# 2. Load scRNASeq dataset
Next, we'll load the A549 dataset, which is the lung cancer cell line investigated in GSE147405. 
```{r, include=FALSE}
# Linux
#a549 = readRDS("~/Data/scRNASeq/GSE147405/A549_TGFB1.rds")

# Dell
magic_datapath = "D:/Analysis/EMT/a549_magic.xlsx"
a549_zscore = read_excel(magic_datapath, sheet="Zscore")
a549_pvalue = read_excel(magic_datapath, sheet="pvalue")
```

Load Seurat object
```{r}
seuratobj_path = "D:/Data/scRNASeq/GSE147405/A549/A549_TGFB1.rds"
a549_tgfb = readRDS(seuratobj_path)
a549_tgfb = subset(a549_tgfb, subset = Time==c("0d", "8h", "3d", "1d", "7d"))
```

# 3. Map dataset 

## Map to time and clusters
```{r}
ids = a549_zscore[, 1]
ids$time = a549_tgfb@meta.data$Time
ids$cluster = a549_tgfb@active.ident
```

## Preprocess the Z-score and p-value
Now let's merge the MAGIC imputed data with the gene identifiers.
```{r}
a549_zscore = data.frame(a549_zscore)
a549_pvalue = data.frame(a549_pvalue)

rownames(a549_zscore) = a549_zscore[, 1]
rownames(a549_pvalue) = a549_pvalue[, 1]

a549_zscore[, 1] = NULL
a549_pvalue[, 1] = NULL

a549_zscore = t(a549_zscore)
a549_pvalue = t(a549_pvalue)

colnames(a549_zscore) = ids$time
colnames(a549_pvalue) = ids$time
```

# 4. Create a gene identifier map
Use BiomaRt to get Ensembl IDs map
```{r}
ensembl = useMart("ensembl", dataset="hsapiens_gene_ensembl")
gene_ids = data.frame(colnames(a549_zscore))
names(gene_ids) = "Gene"
human_ids = getBM(attributes=c('ensembl_gene_id',
                               'entrezgene_id',
                               'hgnc_symbol'), 
                  values=gene_ids,
                  mart=ensembl)
mapped_ids = merge(gene_ids, human_ids, 
                   by.x='Gene', by.y='hgnc_symbol')
```

# 5. Map to Ensembl identifiers
Merge the data with gene identifiers
```{r}
merged_zscore = merge(mapped_ids, a549_zscore, 
                              by.x='Gene', by.y="row.names")
merged_pvalue = merge(mapped_ids, a549_pvalue, 
                              by.x='Gene', by.y="row.names")
```

# 6. Save Ensembl data
```{r}
savepath = "D:/Analysis/EMT/mapped_magic_a549.xlsx"
writexl::write_xlsx(
  x=list(zscore = merged_zscore,
         pvalue = merged_pvalue),
  path=savepath)
```