---
title: "Create expression data matrix"
author: Scott Campit
output: html_notebook
---

## Summary
This notebook concatenates the expression datasets for visualization. The final data matrix will be the effect size for a given set of time points that overlap across all datasets. This will be used to construct a heatmap.

```{r}
rm(list=ls())
```

Load libraries
```{r}
# Google API
library(googledrive)
library(googlesheets4)

# Essential data manipulation libraries
#install.packages(c("BiocManager", "tidyverse", "readxl", 'data.table', 'openxlsx'))
library(data.table)
library(openxlsx)
library(readxl)

# Bioinformatics libraries
#BiocManager::install(c("edgeR", "limma", "GEOquery", "affy",
#                       "hgu133plus2.db", "org.Hs.eg.db"), 
#                     dependencies=TRUE,
#                     update=TRUE)


library(hgu133plus2.db)
library(org.Hs.eg.db)
library(edgeR)
library(GEOquery)
library(affy)
```

## 1. Load Data 
This section will take the expression data and try to overlay it with the predicted fluxes.

### A. Bulk Proteomics Data
Bulk proteomics data was taken from the following sources:
  * Lu, C., Sidoli, S., Kulej, K., Ross, K., Wu, C. H., & Garcia, B. A. (2019). Coordination between TGF-β cellular signaling and epigenetic regulation during epithelial to mesenchymal transition. Epigenetics & chromatin, 12(1), 1-17.
  * Sartor, M. A., Mahavisno, V., Keshamouni, V. G., Cavalcoli, J., Wright, Z., Karnovsky, A., ... & Omenn, G. S. (2010). ConceptGen: a gene set enrichment and gene set relation mapping tool. Bioinformatics, 26(4), 456-463.
  
Note we had to contact the lead author for the Keshamouni proteomics data.

```{r}
# Garcia Proteomics
url = "https://docs.google.com/spreadsheets/d/1LDRQKmoFR4i5swABQ00lm5v7JqubIduFUTTV_YbUGsU/edit?usp=sharing"
garcia_proteomics = read_sheet(url)

# Keshamouni Proteomics
url = "https://docs.google.com/spreadsheets/d/1Gx1b2uvqBRht_iC5nhu0dpcDKX52ymVJjbhFHANEYIg/edit?usp=sharing"
kesh_proteomics = read_sheet(url)
```

### B. Bulk Transcriptomics Data
Bulk transcriptomics data was taken from the following sources:
  * (GSE17708)[https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE17708]
  * (GSE17518)[https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE17518]

```{r}
# Keshamouni Transcriptomics
kesh_trans = getGEO("GSE17708", GSEMatrix = TRUE)
#file.remove("C:\Users\Scott\AppData\Local\Temp\Rtmp0sdK0p/GPL570.soft")

# Thannickal Transcriptomics
than_trans = getGEO("GSE17518", GSEMatrix = TRUE)
#file.remove("C:\Users\Scott\AppData\Local\Temp\Rtmp0sdK0p/GPL96.soft")
```

### C. Single-Cell Transcriptomics Data
Finally, single-cell transcriptomics data was taken from Cook & Vanderhyden, 2020.
  * Cook, D. P., & Vanderhyden, B. C. (2020). Context specificity of the EMT transcriptional response. Nature communications, 11(1), 1-9.
  
```{r}
# Cook & Vanderhyden Transcriptomics


```
