---
title: "Create expression data matrix"
author: Scott Campit
output: html_notebook
---

## Summary
This notebook concatenates the expression datasets for visualization. The final data matrix will be the effect size for a given set of time points that overlap across all datasets. This will be used to construct a heatmap.

```{r}
rm(list=ls())
```

Load libraries
```{r}
# Google API
library(googledrive)
library(googlesheets4)

# Essential data manipulation libraries
#install.packages(c("BiocManager", "tidyverse", "readxl", 'data.table', 'openxlsx'))
library(data.table)
library(openxlsx)
library(readxl)

# Bioinformatics libraries
#BiocManager::install(c("edgeR", "limma", "GEOquery", "affy",
#                       "hgu133plus2.db", "org.Hs.eg.db"), 
#                     dependencies=TRUE,
#                     update=TRUE)


library(hgu133plus2.db)
library(org.Hs.eg.db)
library(edgeR)
library(GEOquery)
library(affy)
```

## 1. Load Data 
This section will take the expression data and try to overlay it with the predicted fluxes.

### A. Bulk Proteomics Data
Bulk proteomics data was taken from the following sources:
  * Lu, C., Sidoli, S., Kulej, K., Ross, K., Wu, C. H., & Garcia, B. A. (2019). Coordination between TGF-β cellular signaling and epigenetic regulation during epithelial to mesenchymal transition. Epigenetics & chromatin, 12(1), 1-17.
  * Sartor, M. A., Mahavisno, V., Keshamouni, V. G., Cavalcoli, J., Wright, Z., Karnovsky, A., ... & Omenn, G. S. (2010). ConceptGen: a gene set enrichment and gene set relation mapping tool. Bioinformatics, 26(4), 456-463.
  
Note we had to contact the lead author for the Keshamouni proteomics data.

```{r}
# Garcia Proteomics
url = "https://docs.google.com/spreadsheets/d/1LDRQKmoFR4i5swABQ00lm5v7JqubIduFUTTV_YbUGsU/edit#gid=1182791926"
garcia_proteomics_effect = read_sheet(url)
url = "https://docs.google.com/spreadsheets/d/1LDRQKmoFR4i5swABQ00lm5v7JqubIduFUTTV_YbUGsU/edit#gid=559844629"
garcia_proteomics_pvalue = read_sheet(url)

# Keshamouni Proteomics
datapath = "D:/Chandrasekaran/Projects/EMT/Data/Proteomics/Keshamouni/raw/Proteomics2010.xlsx"
kesh_proteomics = read_excel(datapath, sheet='Protein List')
```

### B. Bulk Transcriptomics Data
Bulk transcriptomics data was taken from the following sources:
  * (GSE17708)[https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE17708]
  * (GSE17518)[https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE17518]

```{r}
# Keshamouni Transcriptomics
kesh_trans = "D:/Chandrasekaran/Projects/EMT/Data/RNASeq/GSE17708/GSE17708_Keshamouni_TGFB1_logs.xls"
kesh_tran_effect = read_excel(kesh_trans, sheet="FC")
kesh_tran_pvalue = read_excel(kesh_trans, sheet="PValue")

# Thannickal et al Transcriptomics
than_tran = "D:/Chandrasekaran/Projects/EMT/Data/RNASeq/GSE17518/GSE17518_DGE.xlsx"
than_tran = read_excel(than_tran)
```

### C. Single-Cell Transcriptomics Data
Finally, single-cell transcriptomics data was taken from Cook & Vanderhyden, 2020.
  * Cook, D. P., & Vanderhyden, B. C. (2020). Context specificity of the EMT transcriptional response. Nature communications, 11(1), 1-9.
  
```{r}
# Linux Path
#filename = '~/Data/RNASeq/GSE17518/GSE17518_DGE.xlsx'

# Windows -> Dell
#filename = "D:/Chandrasekaran/Projects/EMT/Data/scRNASeq/GSE147405/A549/dge_a549_tgfb1.csv"

filename = 'D:/Chandrasekaran/Projects/EMT/Data/scRNASeq/GSE147405/A549/dge_a549_tgfb1.csv'
cook_vand_data = read.csv(file=filename)
cook_vand_data$X = NULL
```

## 2. Map all of the data to a single dataframe
We need to convert all data (mapped to symbols) to a common index. We should take the outer, so that we'll be able to get all genes.  

### A. Create a function that will perform this mapping

#### i. Map between mouse and human
This is the function that maps genes between two different organisms.
```{r, results='hide'}
# Make a function that maps genes using two reference databases
convertGeneList = function(gene_obj, db1, db2){
  # convertGeneList converts MGI symbols to human identifiers.
  # INPUTS:
  #   * gene_obj: a Seurat object
  #   * db1: A biomart database. This function is specific to the mouse database.
  #   * db2: A biomart database. This function is specific to the human database.
  # OUTPUT:
  #   * mapped_gene_obj: A dataframe with the mapped gene object.   
  orthologs = getLDS(attributes=c("mgi_symbol"), 
                     filters="mgi_symbol", 
                     values=gene_obj$`Gene Symbol`, 
                     mart=db2, 
                     attributesL=c("hgnc_symbol", 
                                   "entrezgene_id", 
                                   "ensembl_gene_id"), 
                     martL=db1, 
                     uniqueRows=TRUE)
  
  mapped_gene_obj = merge(orthologs, 
                          gene_obj, 
                          by.x='MGI.symbol', 
                          by.y='Gene Symbol')
  return(mapped_gene_obj)
}
```

#### ii. Map between identifier types
This function uses a database to map gene/protein IDs to a common identifier
```{r}
geneIDMapping = function(df, geneList, query, db){
  #
  #
  #
  #
  
  map = select(db, 
               keys=geneList, 
               columns = query
               )
  
  # Join map and data
  new_df = cbind(geneList, df)
  mapped_lst = merge(map, new_df, by.x="PROBEID", by.y="probes")
  
  maplist = list("Map1"=map1, "Map2"=map2)
  return(maplist)
}
```

### B. Map datasets 

#### i. Load Single-Cell Data 
```{r}
# Cook and Vanderhyden
map = select(hgu133plus2.db, 
             keys=cook_vand_data$Gene, 
             columns = c('ENTREZID', 
                         'SYMBOL'),
             keytype="SYMBOL")
cookvand = merge(map, cook_vand_data, by.x="SYMBOL", by.y="Gene")
```
#### ii. Load Proteomics datasets
```{r}
# Keshamouni Proteomics
map = select(hgu133plus2.db, 
             keys=kesh_proteomics$REFSEQ, 
             columns = c('ENTREZID', 
                         'SYMBOL',
                         'REFSEQ'),
             keytype="REFSEQ")
kesh_prot = merge(map, kesh_proteomics, by.x="REFSEQ", by.y="REFSEQ")
```

```{r}
# Garcia Proteomics Data
library(biomaRt)
human = useMart("ensembl", dataset = "hsapiens_gene_ensembl")
mouse = useMart("ensembl", dataset = "mmusculus_gene_ensembl")
garcia_prot = convertGeneList(garcia_proteomics_effect, human, mouse)
```

#### Create a universal map for Entrez IDs and Gene Symbols

First, replace all column names to be consistent.
```{r}
colnames(garcia_prot)[which(names(garcia_prot) == "HGNC.symbol")] = "SYMBOL"
colnames(garcia_prot)[which(names(garcia_prot) == "NCBI.gene..formerly.Entrezgene..ID")] = "ENTREZID"

colnames(kesh_tran_effect)[which(names(kesh_tran_effect) == "Gene Symbol")] = "SYMBOL"
colnames(kesh_tran_effect)[which(names(kesh_tran_effect) == "Entrez Gene")] = "ENTREZID"
```

Then ensure the data types are the same for each column.
```{r}
garcia_prot = garcia_prot %>% 
  dplyr::mutate(
    SYMBOL = as.character(SYMBOL),
    ENTREZID = as.character(ENTREZID)
  )

kesh_prot = kesh_prot %>% 
  dplyr::mutate(
    SYMBOL = as.character(SYMBOL),
    ENTREZID = as.character(ENTREZID)
  )

kesh_tran_effect = kesh_tran_effect %>% 
  dplyr::mutate(
    SYMBOL = as.character(SYMBOL),
    ENTREZID = as.character(ENTREZID)
  )

than_tran = than_tran %>% 
  dplyr::mutate(
    SYMBOL = as.character(SYMBOL),
    ENTREZID = as.character(ENTREZID)
  )

cookvand = cookvand %>% 
  dplyr::mutate(
    SYMBOL = as.character(SYMBOL),
    ENTREZID = as.character(ENTREZID)
  )

```

Finally, combine all identifiers into a single dataframe.
```{r}
universal_map = dplyr::bind_rows(kesh_prot[,        c("SYMBOL", "ENTREZID")],
                                 garcia_prot[,      c("SYMBOL", "ENTREZID")],
                                 kesh_tran_effect[, c("SYMBOL", "ENTREZID")],
                                 than_tran[,        c("SYMBOL", "ENTREZID")],
                                 cookvand[,         c("SYMBOL", "ENTREZID")])
universal_map = dplyr::distinct(universal_map)
```

Now, we'll extract the relevant columns that we need.
```{r}
colnames(cookvand)
kesh_prot        = kesh_prot[, c("SYMBOL", "ENTREZID", 
                                 "Log2 ratio")]
garcia_prot      = garcia_prot[, c("SYMBOL", "ENTREZID", 
                                   "5 min/0 min", "60 min/0min", "1 day/0 min", "2 days/0 min")]
kesh_tran_effect = kesh_tran_effect[, c("SYMBOL", "ENTREZID", 
                                        "0.5 vs 0", "1 vs 0", "2 vs 0", "4 vs 0", "8 vs 0", "16 vs 0", "24 vs 0", "72 vs 0")]
than_tran        = than_tran[, c("SYMBOL", "ENTREZID", "t")]
cookvand         = cookvand[, c("SYMBOL", "ENTREZID", "Spline_1", "Spline_2", "Spline_3")]
```

Then ensure the data types are the same for each column.
```{r}
colnames(kesh_prot)[which(names(kesh_prot) == "Log2 ratio")] = "Keshamouni_72hrs"

colnames(garcia_prot)[which(names(garcia_prot) == "60 min/0min")] = "Garcia_1hrs"
colnames(garcia_prot)[which(names(garcia_prot) == "1 day/0 min")] = "Garcia_24hrs"
colnames(garcia_prot)[which(names(garcia_prot) == "2 days/0 min")] = "Garcia_48hrs"

colnames(kesh_tran_effect)[which(names(kesh_tran_effect) == "1 vs 0")] = "GSE17708_1hrs"
colnames(kesh_tran_effect)[which(names(kesh_tran_effect) == "8 vs 0")] = "GSE17708_8hrs"
colnames(kesh_tran_effect)[which(names(kesh_tran_effect) == "24 vs 0")] = "GSE17708_24hrs"
colnames(kesh_tran_effect)[which(names(kesh_tran_effect) == "72 vs 0")] = "GSE17708_72hrs"

colnames(than_tran)[which(names(than_tran) == "t")] = "GSE17518_72hrs"

colnames(cookvand)[which(names(cookvand) == "Spline_1")] = "GSE147405_8hrs"
colnames(cookvand)[which(names(cookvand) == "Spline_2")] = "GSE147405_24hrs"
colnames(cookvand)[which(names(cookvand) == "Spline_3")] = "GSE147405_72hrs"
```

```{r}
data = list(kesh_prot[, c("SYMBOL", "ENTREZID", 
                          "Keshamouni_72hrs")],
            garcia_prot[, c("ENTREZID", 
                            "Garcia_1hrs", "Garcia_24hrs", "Garcia_48hrs")], 
            kesh_tran_effect[, c("ENTREZID", 
                                 "GSE17708_1hrs", "GSE17708_8hrs", "GSE17708_24hrs", "GSE17708_72hrs")], 
            than_tran[, c("ENTREZID", 
                          "GSE17518_72hrs")], 
            cookvand[, c("ENTREZID", 
                         "GSE147405_8hrs", "GSE147405_24hrs", "GSE147405_72hrs")]) 
```

```{r}
kesh_prot = kesh_prot[, c("ENTREZID", 
                          "Keshamouni_72hrs")]

garcia_prot = garcia_prot[, c("ENTREZID", 
                "Garcia_1hrs", "Garcia_24hrs", "Garcia_48hrs")]

kesh_tran_effect = kesh_tran_effect[, c("ENTREZID", 
                     "GSE17708_1hrs", "GSE17708_8hrs", "GSE17708_24hrs", "GSE17708_72hrs")]

than_tran = than_tran[, c("ENTREZID", 
                          "GSE17518_72hrs")]

cookvand = cookvand[, c("ENTREZID", 
                         "GSE147405_8hrs", "GSE147405_24hrs", "GSE147405_72hrs")]
```

```{r}
kesh_tran_effect = aggregate(. ~ENTREZID, data=kesh_tran_effect, median, na.rm=TRUE)
garcia_prot = aggregate(. ~ENTREZID, data=garcia_prot, median, na.rm=TRUE)
kesh_prot = aggregate(. ~ENTREZID, data=kesh_prot, median, na.rm=TRUE)
than_tran = aggregate(. ~ENTREZID, data=than_tran, median, na.rm=TRUE)
cookvand = aggregate(. ~ENTREZID, data=cookvand, median, na.rm=TRUE)
```

```{r}
data = merge(universal_map, kesh_tran_effect, by="ENTREZID", all= TRUE)
data = merge(data, garcia_prot, by="ENTREZID", all= TRUE)
data = merge(kesh_prot, data, by="ENTREZID", all= TRUE)
data = merge(data, than_tran, by="ENTREZID",  all= TRUE)
data = merge(data, cookvand, by="ENTREZID",  all= TRUE)
```

```{r}
geneset_path = "D:/Chandrasekaran/Projects/EMT/Analysis/data/glycolysis_geneset.txt"
library(readr)
geneset = read_csv(geneset_path)
glycolysis_genes = merge(data, geneset, by.x='SYMBOL', by.y='KEGG_GLYCOLYSIS_GLUCONEOGENESIS')
```

```{r}
# Split by hours
early_data = glycolysis_genes[, c("GSE17708_1hrs", "Garcia_1hrs")]
mid1_data  = glycolysis_genes[, c("GSE17708_8hrs", "GSE147405_8hrs")]
mid2_data  = glycolysis_genes[, c("GSE17708_24hrs", "Garcia_24hrs", "GSE147405_24hrs")]
late_data  = glycolysis_genes[, c("GSE17708_72hrs", "GSE17518_72hrs", "Garcia_48hrs", 
                       "Keshamouni_72hrs", "GSE147405_72hrs")]
```

```{r}
#BiocManager::install("ComplexHeatmap")
#install.packages('circlize')
library(ComplexHeatmap)
library(circlize)

h1_col = HeatmapAnnotation(h1=anno_block(gp=gpar(fill="#CAC4CE"),
                                      labels=c("1Hr"), 
                                      labels_gp=gpar(col="black", 
                                                     fontsize=10)))
h2_col = HeatmapAnnotation(h2=anno_block(gp=gpar(fill="#8D86C9"),
                                      labels=c("8Hrs"), 
                                      labels_gp=gpar(col="white", 
                                                     fontsize=10)))
h3_col = HeatmapAnnotation(h3=anno_block(gp=gpar(fill="#725AC1"),
                                      labels=c("24Hrs"), 
                                      labels_gp=gpar(col="white", 
                                                     fontsize=10)))
h4_col = HeatmapAnnotation(h4=anno_block(gp=gpar(fill="#242038"),
                                      labels=c("48-72Hrs"), 
                                      labels_gp=gpar(col="white", 
                                                     fontsize=10)))
```

```{r, fig.width=20, fig.height=20, fig.align='right', dpi=100}
col_fun = colorRamp2(c(-3, 0, 3), 
                     c("#2B9EB3", "Gray", "#F8333C"))
h1 = Heatmap(row_labels=glycolysis_genes$SYMBOL,
             row_names_side="left",
             early_data, 
             col=col_fun, 
             cluster_columns=FALSE, 
               rect_gp = gpar(col="black", 
                              lwd=1),
             cluster_rows=FALSE,
             show_heatmap_legend=FALSE,
             bottom_annotation=h1_col, 
             width = ncol(early_data)*unit(5, "mm"), 
             height = nrow(early_data)*unit(5, "mm"))
h2 = Heatmap(mid1_data, 
             col=col_fun, 
             cluster_columns=FALSE, 
               rect_gp = gpar(col="black", 
                              lwd=1),
             cluster_rows=FALSE,
             show_heatmap_legend=FALSE,
             bottom_annotation=h2_col, 
             width = ncol(mid1_data)*unit(5, "mm"), 
             height = nrow(mid1_data)*unit(5, "mm"))
h3 = Heatmap(mid2_data, 
             col=col_fun, 
             cluster_columns=FALSE, 
               rect_gp = gpar(col="black", 
                              lwd=1),
             
             cluster_rows=FALSE,
             show_heatmap_legend=FALSE,
             bottom_annotation=h3_col, 
             width = ncol(mid2_data)*unit(5, "mm"), 
             height = nrow(mid2_data)*unit(5, "mm"))
h4 = Heatmap(late_data, 
             col=col_fun, 
             cluster_columns=FALSE, 
               rect_gp = gpar(col="black", 
                              lwd=1),
             cluster_rows=FALSE,
             show_heatmap_legend=TRUE,
             bottom_annotation=h4_col,
             width = ncol(late_data)*unit(5, "mm"), 
             height = nrow(late_data)*unit(5, "mm"),
             heatmap_legend_param=list(title="Expression"))

final_figure = h1+h2+h3+h4
final_figure
```